% !TeX spellcheck = en_GB
% !TeX encoding = UTF-8
% !TeX root = ../report.tex

\chapter{Communication}
\label{chp:Communication}

To handle the communication between all devices, a fast and easy to implement data exchange method was required. RIMO, as well as the other manufacturers of our components made use of the standardized industrial application CANOpen. Because CANOpen is based on CAN, we will first describe the CAN bus protocol.

\section{CAN Basics}


CAN stands for Control Area Network and consists of two main layers, namely the physical layer and the data link layer. CAN was developed in 1986 and is used for data exchange between different stations in a network, based on serial communication. Messages are received and transmitted via broadcasting, making every message available for all of the connected stations. 
%importance and pros cons of can

The rise in electronification in many parts of the industry called for simpler and more efficient means of communication. Extensive wiring still resulted in rather limited data exchange. A way out of this was presented by serial bit data exchange and connecting up all electronic control units to a single bus. Depending on the bus length, CAN offers up to 1 Mbit/s of data rate, while remaining robust and reliable, even in a noisy environment.

The physical layer consists of a twisted pair of wires, which can be shielded if required. The value of a bit was determined by the difference in voltage of the two wires. If the voltages are the same, the bit is recessive. If the difference is higher than 0.9 V, the bit is dominant. As both wires are affected by the same electromagnetic disturbances, their difference in voltage will not vary. The data link is therefore immune to electromagnetic disturbances. By twisting the wires, the magnetic field generated will also be reduced significantly.

%graphic of can network
%elaborate on physical layer and simple message, 60 ohm when measured as both ends parallel
%baud rate

To reduce reflections in the data cables with rates higher than 125kbit/s, it is recommended to terminate the ends of the communication lines with termination resistors with 120 Ohm. 

Each CAN message contains the following structure.
%picture and example of can message

Especially the role of the identifier bit is important, as it handles and represents the message priority.

\pagebreak

\section{CANOpen}

CANOpen is a higher-layer protocol based on the aforementioned CAN. In the following, we will provide information on the project-relevant aspects of CANOpen.
All real-time data is exchanged via process data objects (PDO). The adress of each individual PDO can be found in a standardized table by CAN in Automation (CiA). PDO's contain data from a single or different objects from the object dictionary. This maps each bit of the data section of a PDO to a certain object.

To change PDO mapping or customise other parameters stored in the object dictionary, service data objects (SDO) are used. These however do not transmit and receive real-time data, but are only used for service purposes. They usually contain the index, subindex and the new value of the respective object.

PDO as well as SDO communication is split into receive and transmit messages.
 For PDO communication the receive PDO contains the desired command for a node and the Transmit PDO is composed of the actual state or other valuable information measured on node itself.

In SDO communication a receive-SDO is either used to change a value in the object dictionary or to request the contents of an entry in the object dictionary. The receive-SDO is always answered by a corresponding transmit-PDO either acknoledging a change in the object dictionary or serving the requested data.

So for example, the mapping of the ACD's RPDO1 looks like the following.

\begin{tabular}{|c|c|c|c|}
	\hline 
	Name & Index & Subindex & Bit \\ 
	\hline 
	&  &  &  \\ 
	\hline 
	Command Word & 0x2000 & 1 & 16 \\ 
	\hline 
	Command Speed & 0x2000 & 2 & 16 \\ 
	\hline 
	Command Acceleration & 0x2000 & 5 & 8 \\ 
	\hline 
	Command Deceleration & 0x2000 & 6 & 8 \\ 
	\hline 
\end{tabular} 
\\
\\
\textit{Table 1, RPDO1 of the ACD 4805}

So in order to set a desired speed on the right (Node 6) motor controller of around 900 rpm, the following message had to be sent.

\begin{tabular}{|c|c|c|c|c|c|c|}
	\hline 
	CAN-ID & \multicolumn{6}{c|}{DATA} \\ 
	\hline 
	0x206 & 09 & 00 & 84 & 03 & 00 & 00 \\ 
	\hline 
\end{tabular} 

Take note that to translate the desired target input into a bytewise message
the little endian rule is most often utilized and the hex format is used to display the contents of each byte. According to the little endian rule the byte containing the higher bit values comes second in transmission. 
The Command speed of the message above is split into two hex numbers yielding the underlying number 384h.

Each CAN device has a personal node, ranging from 1 to 127. This ensures that every message reaches it's corresponding device.
% example of pdo can message


% example of sdo can message
%keywords
%transmission type, can id, sdo, pdo, NMT
There are various transmission types avaible in CANOpen for PDOs. Some are manufacturer specific, therefore only the types relevant for our project will shortly be described.

\begin{itemize}

\item Transmission type 1-240 (cyclic synchronous)
\begin{itemize}
	\item This transmission type relies on sync messages. After every nth sync messages, the PDO transmits it's data. So for example, if the transmission type is 6, the PDO will send after every 6th sync message.
	
	For the Linmot, we used the transmission type one. So in order to gain information about the motor, we had to send sync messages.
	
\end{itemize}

\item {Transmission type 254/255 (asynchronous)}
\begin{itemize}
	\item These transmission types are event-triggered, where the event is manufacturer specific for 254 and for 255 defined in the CANOpen device profile.
	
	For the ACD 4805, tranmission type 254 was listed as unpack the data immediately. This was the setting used in the ACD and in the Mircoautobox.
	
	For the LinMot, transmission type 254 was event triggered, with an adjustable internal timer which triggers the event.
\end{itemize}

\end{itemize}


%busheavy, buslight, busoff, correct baudrate etc.



The standard table of CAN-Id's can be found below.

\begin{tabular}{|c|c|c|}
	\hline 
	\textbf{COB} & \textbf{Function Code} & \textbf{Resulting CAN-ID} \\ 
	\hline 
	NMT &  0000$_{b}$  & $0 (000_{h})$  \\ 
	\hline 
	CODE & $0001_{b}$ &  $128 (080_{h})$\\ 
	\hline 
	TIME & $0010_{b}$ &  $256 (100_{h})$\\ 
	\hline 
	EMCY & $0001_{b}$ &   $129 (081_{h}) $–  255 (0FF$_{h})$ \\ 
	\hline 
	PDO1(tx) & $0011_{b}$ &  $385 (181_{h}) $ – 511 (1FF$_{h})$\\ 
	\hline 
	PDO1(rx) & $0100_{b}$ &  $513 (201_{h}) $ –  639 (27F$_{h})$\\ 
	\hline 
	PDO2(tx) & $0101_{b}$ &  $641 (281_{h}) $ –  767 (2FF$_{h})$\\ 
	\hline 
	PDO2(rx) & $0110_{b}$ &  $769 (301_{h})$  –  895 (37F$_{h})$\\ 
	\hline 
	PDO3(tx) & $0111_{b}$ &  $897 (381_{h}) $ – 1023 (3FF$_{h})$\\ 
	\hline 
	PDO3(rx) & $1000_{b}$ &  $1025 (401_{h}) $ –  1151 (47F$_{h})$\\ 
	\hline 
	PDO4(tx) & $1001_{b}$ &  $1153 (481_{h}) $ –  1279 (4FF$_{h})$\\ 
	\hline 
	PDO4(rx) & $1010_{b}$ & $1281 (501_{h}) $ –  1407 (57F$_{h})$\\ 
	\hline 
	SDO(tx) & $1011_{b}$ & $ 1409 (581_{h}) $ –  1535 (5FF$_{h})$\\ 
	\hline 
	SDO(rx) & $1100_{b}$ & $1537 (601_{h}) $ –  1663 (67F$_{h})$\\ 
	\hline 
	NMT error control &$ 1110_{b} $& $ 1793 (701_{h}) $ –  1919 (77F$_{h})$\\ 
	\hline 
\end{tabular} 
\\
\\
\textit{Table 2, COBs and there corresponding CAN-IDs}
%adjust table number







%brief explaination on additional features, their advantages and how they benefited us.
\subsection{Implementation}
%explain what commands we used, object dictionary and relevant commands/messages.


\subsubsection{ACD}

%write about RPDO1 RPDO2 conflict, our assumptions what if couldve been, namely phyiscal connection, sync messages, overload, messages overlapping etc. solved by disabling rpdo2

In order to communicate with the ACD, the CAN Node-Id of each ACD had to be determined. One of Rimo's pdf files, namely "Setting up ACD Controller \& Connection Diagram", serves exactly this purpose. 
%refer to appendix witht file.
To find the respective node-id's, we had to check if either pin 12 (DI5) or pin 20 (DI6) was connected to any other pins on the ACD 4805 K1 connector. In our case, pin 12 of the right ACD was connected to pin 1, making it node 6. Pin 12 of the left ACD was not connected and therefore making it node 5. Afterwards our findings were confirmed when we connected the go-kart to our CAN bus. With the help of a CAN-USB adapter, we were able to receive and send messages, and after a while controlling the kart.
Because the go-kart does not communicate via CAN by default but only for service and remote control purposes, we had to put all nodes into operational mode by sending a NMT start messages to all nodes, namely node 5 and 6.
The ID of a NMT message is 0x000, to start the nodes the instruction code 0x01 had to be used. To reach all nodes simultaneously, the node adress needed to be 0x0. 

Issues with CANOpen and dSpace. 

When we tried to control the ACD via CANOpen with the Microautobox, the communication presented itself to be an issue. The wheels were not turning smooth, but rather interrupted from time to time. It was clear that something had to be disrupting the communication.
With the help of a CAN-USB adapter, we were able to monitor the CAN bus. After checking all physical layers, we came to the conclusion that it must be a software problem. Our first approach was to change synchronisation times. This however did not have any effect, neither did changing the baud rate or adjusting the step size for Matlab's solver.
A smooth motion of the motors could be realized by sending the Receive PDO manually via the Pcan software. The problem turned out to arise from the second Receive PDO. While the first RPDO determines the target speed for the speed controller the second RPDO provides the possibility to control the motor with target torque. By default, the use of one of the mentioned does not disable the function of the other. This leads to critical failure  causing improper motion if both RPDOs are sent simultaneously. The Motor tries to achieve both, target speed and target torgue ending up rapidly switching the motor current. The problem could be solved by completely removing the second RPDO from the Matlab model. The CANopen blocks do not allow for deactivation of the individual PDO's transmission in real time.


We established connection between MABX and a receiver, in our case a laptop connected to the MABX CAN via CAN USB.


Two's complement is convenient way to store integers, such that adding and subtracting with negative numbers becomes very easy. This was used on the ACD 4805, where certain values, such as the rotational speed of the wheels, can be negative. 
The basic principles of two's complement are the following.

- Zero is represented by all 0's. 
e.g. 0 0 0 0 = 0

- The maximum positive integer is $ 2^{number of bits-1}-1 $.
So for 4 bits, the biggest integer is therefore 0 1 1 1 = 7, and not 1 1 1 1 = 15 as in the standard notation.

- if the integer is negative, 1's and 0's switch roles, starting from all one's, e.g. 1 1 1 1 = -1. This increases the range for negative numbers by one.

So for 4 bits it looks as follows.

\begin{tabular}{llll}
0 0 0 0 = 0 & 0 1 0 0 = 4 & 1 1 1 1 = -1 & 1 0 1 1 = -5\\
0 0 0 1 = 1 & 0 1 0 1 = 5 & 1 1 1 0 = -2 & 1 0 1 0 = -6\\
0 0 1 0 = 2 & 0 1 1 0 = 6 & 1 1 0 1 = -3 & 1 0 0 1 = -7\\
0 0 1 1 = 3 & 0 1 1 1 = 7 & 1 1 0 0 = -4 & 1 0 0 0 = -8
\end{tabular}

So an easy way to find the negative integer of a positive integer is to convert the desired decimal number to binary, inverting all 0's and 1's and then adding 1. A more hands on approach is to again convert to binary, starting from the right to find the first 1 and inverting all bits to the left of it.

So in order to have a rotational speed of -500 revolutions per minute, the following steps have to be taken for a signed 16 bit integer.

500 = 0 0 0 0 \: 0 0 0 1 \: 1 1 1 1 \: 0 1 0 0

Now invert all bits.

1 1 1 1 \: 1 1 1 0 \: 0 0 0 0 \: 1 0 1 1

And add 1

1 1 1 1 \: 1 1 1 0 \: 0 0 0 0 \: 1 1 0 0 = -500

The second method would result in the following steps.

Starting from the right, find first 1.

500 = 0 0 0 0 \: 0 0 0 1 \: 1 1 1 1 \: 0 \textbf{1} 0 0

Invert all consecutive bits to the left of it.

1 1 1 1 \: 1 1 1 0 \: 0 0 0 0 \: 1 \textbf{1} 0 0 = -500

If a signed integer is postive or negative is easy to spot, as it's most significant bit determines if the number is negative or positive. 1 = negative, 0 = positive.

\subsubsection{LinMot}
important settings

For some reason, TPDO3 and TPDO4 would not transmit their data when a sync message was sent. This was tried to solve by setting the intern event timer to around 10 ms and changing the transmission type to 254. However, after a while it would stop transmitting out of a unknown reason. As time was scarce, we simply remapped the PDOs, such that the needed data would be transmitted via TPDO2. 

%talk about failsafe for brake

Another issues presented itself while working on the brake. Previously we used a motion command called VAI go to position 16bit, which takes velocity, acceleration/deceleration and position as input and creates a curve for the linear motor, which results in the motion. However, a new command will only be executed if the last four bits of the motion command header, called command count, has changed. In the easiest way, bit 0 can be toggled. 
To avoid this, a different setting could be utilized, called PV Stream. This uses a constant stream of position and velocity inputs during a fixed streaming period, interpolates and executes the command. While this seemed very intriguing, it's implementation was not possible. For some reason an error arose, stating that our streaming was too slow. Even after checking the period time with PCAN and checking all settings, the issue could not be resolved.
After that, we went back to the prior way of setting the position.

To counteract the need for change in the command count in order for a change in motion we introduced a pulse generator into the Matlab model.
The pulse generators output is an alternating signal either being zero or one. With the signal from the pulse generator added to the command header, the command count changes every halve period of the pulse generator. With a sufficiently low pulse generator period a function similar to the PV stream could be implemented.
%%really
%clock not working as it triggers resets all the time

\section{Matlab Model}

In the following passage, some of the most important components of the matlab model will be explained. The matlab model is the underlying software basis for the C code running on the Microautobox handling the CAN bus and all communication on it.

The physical layer of CAN has been set up in the multimessage controllersetup block. The baudrate has been set to 250 Kb/s. The module together with the board number determine the pins for the CAN connection. Moreover, Various status messages of the CAN bus can be read out from this block.

The CANOpen general setup block generates a matlab block for each node on the CAN bus, sets the transmission type for each message and determines the sync interval for the network. The Sync interval has been set to 2ms and the transmission type has been defined as cyclic synchronous for all nodes. The heartbeat message and nodeguarding have been turned off in this first prototype.

The blocks generated by the general setup block allow for numerous inputs and outputs. Some of the most important functionalities are the following:

\begin{description}
	\item[Sync Block]
	The sync block generates the sync block according to the interval defined earlier. The input turns the sync message on or off.
	\item[Reset] 
	By feeding a rising edge into this input a node can be resetted. 
	\item[NMT Timeout]
	The time the nodes have to restart after a reset. If a node takes longer to restart than stated by this expression an error will occur.
	\item[SDO Trigger]
	To send a SDO message a rising edge must be fed into this port.  
	\item[SDO W and RXPDO Data] 
	The data for PDO respectively SDO transmission can be led into this input. The data is composed of vectors of data type uint8, each integer representing on byte.The data will be transmitted with the corrresponding identifier by default.
	\item[SDO R and TXPDO data] 
	The response to a SDO message and the cyclic transmission of the TXPDO can be read out here. The bytes of all TXPDO are encapsuled in one vector.
	
\end{description}

The following picture shows the multimessage controllersetup block and the node blocks created by the CANOpen general setup block.
%Bilder


