% !TeX spellcheck = en_GB
% !TeX encoding = UTF-8
% !TeX root = ../report.tex

\chapter{Introduction}
\label{chp:Introduction}

%keywords 
%drive by wire, components, idea, motivation, objective
%many different solutions and approaches possible

%context of project (go more into detail)
The purpose of this work was to design a drive-by-wire system for future use in the field of autonomous self driving cars. The idea of the overall project is to have a fleet of autonomous vehicles, mainly to test fleet management algorithms. %correct??
To realize this, an electric Kart will be modified such that all mechanical inputs normally made by a human driving the cart can be made by actuator units. A Kart offers a easily customizable basis since all components are without casing and hence simple to change. This go-kart prototype was made as a proof of concept and for early testing of the software and hardware.

%goal and reason for necessity
Our objective therefore was plan and realise a solution for every critical component, i.e. throttle, brake and steering. In order to test A.I. and the fleet management algorithms, a stable and reliable low level platform needs to be in place. Requirements for the indiviual components in the drive-by-wire system were i.e. precision, speed, reliable communication and flexibility. Planing included steps such as finding the parts, how to build it into the kart, communicating with the part and supplying the right amount of power.
%state of the art (different solutions, challenges,why this work and that not etc.)
Designing a drive-by-wire system poses many challenges, which can be approached in many different ways. Most drive-by-wire vehicles are being conceptualised with safety as their highest priority. This is because those vehicles transport passengers and their well-being is of utmost importance. That is why these system contain many fallback levels, which are rather difficult to realise. Our kart does not contain such fallback levels, simply because its self-driving capabilites are only used on racing tracks and the drive-by-wire system will not be active when a human is behind the wheel. 

%shortly describe chapters and their importance,connection, and specific order.

